---
title: "Getting Started with RestCountries"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with RestCountries}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(countries)
```

# Introduction
The Rest Countries datapackage is a wrapper of the popular [Rest Countries](https://restcountries.eu/) API project for getting information about the world's nations via REST calls. These calls allow users to retrieve all available countries or to retreive a given country's currency, capital city, calling code, region, sub-region, ISO 639-1 language, name, or country code.

The API END points output include country's name, full name, code, list of codes, currency, language, capital city, calling code, region, bloc, geographical coordinate, population and more. 

In this vignette I walk through the process of getting started with the restcountries datapackage. 

# Installing `restcountries`

The restcountries package can be downloaded from Github using the remotes package which allows easy installation of R packages from remote repositories such as Github. 
```
install.packages("remotes")
library(remotes)
remotes::install_github("denironyx/restcountries")
```

# Rest Countries endpoints
The restcountriesr datapackage can be filter by the following.

- Search by All
- Search by Name or Full Name
- Country's Code
- Country's Currency
- Search by Capital City
- Search by Calling Code
- Search by Region

When a user filter using any of the above function the following are the column names that are queried in the result.
 "country_name"         "domain"               "iso2c"                "iso3c"               
 "calling_codes"        "capital"              "alt_spelling"         "region"              
 "sub_region"           "population"           "lat"                  "lon"                 
 "demonym"              "area"                 "gini"                 "timezones"           
 "borders"              "native_name"          "numeric_code"         "currencies"          
 "languages_iso639_1"   "languages_iso639_2"   "languages_name"       "languages_nativeName"
 "flag"                 "regional_blocs"       "cioc"
 
 In this vignette I decided to subset the column and data I will be working with to provide a suitable examples. Here are the columns name I will be working with 
 "country_name"         "capital"               "iso2c"                "iso3c" 
 "calling_codes"        "lat"                   "lon"                  "region" 
 "currencies"           "population"
```
parameter <- c("country_name", "capital", "iso2c", "iso3c", "calling_codes", "lat", "lon", "region", "currencies", "population")
```



## Search by All
 The `rc_all` function described the REST endpoint available that you can use to search for all the countries information.

```{r}
library(countries)
library(dplyr)

df <- rc_all()

##List of columns
parameter <- c("country_name", "capital", "iso2c", "iso3c", "calling_codes", "lat", "lon", "region", "currencies", "population")

# The first 6 rows of the result queried
df %>% 
  select(all_of(parameter)) %>% 
  head()
```

## Search by Name
Search by country name. This function will allow you search countries by their short name, official name,native name or partial name. More information can be found https://en.wikipedia.org/wiki/List_of_countries_and_dependencies_and_their_capitals_in_native_languages
  
```{r}
df <- rc_by_name("Nigeria")

##List of columns
parameter <- c("country_name", "capital", "iso2c", "iso3c", "calling_codes", "lat", "lon", "region", "currencies", "population")

# print the result of the queried data
df %>% 
  select(all_of(parameter)) %>% 
  head()
```

## Search by Country's Code
Country codes are a component of the international telephone numbering plan, and are necessary only when dialing a telephone number to establish a call to another country. Country codes are dialed before the national telephone number. By convention, international telephone numbers are represented by prefixing the country code with a plus sign +, which also indicates to the subscriber that the local international call prefix must first be dialed. https://en.wikipedia.org/wiki/List_of_country_calling_codes

```{r}
df <- rc_by_callingcode("234")

##List of columns
parameter <- c("country_name", "capital", "iso2c", "iso3c", "calling_codes", "lat", "lon", "region", "currencies", "population")

#print the result of the query
df %>% 
  select(all_of(parameter)) %>% 
  head()
```






















